---
title: "UC business Analytics R programming Guide 4"
author: "Koji Mizumura"
date: "`r Sys.Date()`"
output:
  word_document:
    toc: yes
    toc_depth: '4'
  rmdformats::readthedown:
    number_sections: yes
    fig_height: 10
    fig_width: 14
    highlight: kate
    toc_depth: 3
    css: style.css
  html_document:
    number_sections: yes
    section_divs: yes
    theme: readable
    toc: yes
    toc_depth: 4
    toc_float: yes
always_allow_html: yes
---

```{r setup4, include=FALSE}
# Set global knitr chunk options
knitr::opts_chunk$set(
  fig.align = "center",
  # fig.height = 4.5,
  error = FALSE,
  message = FALSE, 
  warning = FALSE, 
  collapse = TRUE,
  cache = TRUE
)
```

```{r}
library(tidyverse)
library(forecast)
library(magrittr)
library(tidymodels)
library(readxl)
```



# Exploring and visualizing time series

## Creating time series objects in R
A time series can be thought of as a vector or matrix of numbers along with some information about what times those numbers were recorded. This information is stored in a ts object in R. In most exercises, you will use time series that are part of existing packages. However, if you want to work with your own data, you need to know how to create a `ts` object in R.

Let's look at an example `usnim_2002` below, containing net interest margins for US banks for the year 2002 (source: FFIEC).

The function `ts()` takes in three arguments:

- `data` is set to everything in usnim_2002 except for the date column; it isn't needed since the ts object will store time information separately.
- `start` is set to the form c(year, period) to indicate the time of the first observation. Here, January corresponds with period 1; likewise, a start date in April would refer to 2, July to 3, and October to 4. Thus, period corresponds to the quarter of the year.
- `frequency` is set to 4 because the data are quarterly.
In this exercise, you will read in some time series data from an xlsx file using read_excel(), a function from the readxl package, and store the data as a ts object. Both the xlsx file and package have been loaded into your workspace.

```{r}
# Read the data from Excel into R
mydata <- read_excel("data/exercise1.xlsx")

# Look at the first few lines of mydata
head(mydata)

# Create a ts object called myts
myts <- ts(mydata[, 2:4], start = c(1981, 1), frequency = 4)
```

## Time series plots

The first step in any data analysis task is to plot the data. Graphs enable you to visualize many features of the data, including patterns, unusual observations, changes over time, and relationships between variables. Just as the type of data determines which forecasting method to use, it also determines which graphs are appropriate.

You can use the `autoplot()` function to produce a time plot of the data with or without facets, or panels that display different subsets of data:

```{r eval=FALSE}
autoplot(usnim_2002, facets = FALSe)
```

The above method is one of the many taught in this course that accepts boolean arguments. Both T and TRUE mean "true", and F and FALSE mean "false", however, T and F can be overwritten in your code. Therefore, you should only rely on TRUE and FALSE to set your indicators for the remainder of the course.

You will use two more functions in this exercise, `which.max()` and `frequency()`.
`which.max()` can be used to identify the smallest index of the maximum value

```{r}
x <- c(4, 5, 5)
which.max(x)
# [1] 2
```

To find the number of observations per unit time, use frequency(). Recall the usnim_2002 data from the previous exercise:

```{r eval=FALSE}
> frequency(usnim_2002)
[1] 4
```

Because this course involves the use of the `forecast` and `ggplot2` packages, they have been loaded into your workspace for you, as well as `myts` from the previous exercise and the following three series (available in the package `forecast`):

- `gold` containing gold prices in US dollars
- `woolyrnq` containing information on the production of woollen yarn in Australia
- `gas` containing Australian gas production

```{r}
# Plot the data with facetting
autoplot(myts, facets = TRUE)

# Plot the data without facetting
autoplot(myts, facets = FALSE)

# Plot the three series
autoplot(gold)
autoplot(woolyrnq)
autoplot(gas)

# Find the outlier in the gold series
goldoutlier <- which.max(gold)

# Look at the seasonal frequencies of the three series
frequency(gold)
frequency(woolyrnq)
frequency(gas)
```

## Seasonal plots

Along with time plots, there are other useful ways of plotting data to emphasize seasonal patterns and show changes in these patterns over time.

A seasonal plot is similar to a time plot except that the data are plotted against the individual “seasons” in which the data were observed. You can create one using the `ggseasonplot()` function the same way you do with `autoplot()`.

An interesting variant of a season plot uses polar coordinates, where the time axis is circular rather than horizontal; to make one, simply add a polar argument and set it to `TRUE`.A subseries plot comprises mini time plots for each season. Here, the mean for each season is shown as a blue horizontal line.One way of splitting a time series is by using the `window()` function, which extracts a subset from the object x observed between the times start and end.

```{r}
# Load the fpp2 package
library(fpp2)

# Create plots of the a10 data
autoplot(a10)
ggseasonplot(a10)

# Produce a polar coordinate season plot for the a10 data
ggseasonplot(a10, polar = TRUE)

# Restrict the ausbeer data to start in 1992
beer <- window(ausbeer, start = 1992)

# Make plots of the beer data
autoplot(beer)
ggsubseriesplot(beer)
```

## Trends, seasonality, and cyclicity



# Benchmark methods and forecast accuracy



# Exponential smoothing
# Forecasting with ARIMA models
# Advanced methods


